### 2021-01-12 written/eLifeRevision/Bittner_eLife2020.tex
2D lds up-to-date

EPI optimizes the weights and biases $\bm{\theta}$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
Such convergence is evaluated with a hypothesis test that the means and variances of each emergent property statistic are not different than their constrained values (see Section \ref{methods_AL_opt}). 
Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).

In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $f(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
A key advantage of EPI over established Bayesian inference is that the predictions made my the inferred distribution are constrained to produce the specified emergent property.
%Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
Equipped with this method, we may examine structure in posterior distributions or make comparisons between posteriors conditioned at different levels of the same emergent property statistic.
In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using it to investigate and produce novel insights into two prominent models in neuroscience. 
Subsequently in Section \ref{results_LRRNN}, we show EPI's superiority in parameter scalability and fidelity of the posterior predictive distribution by conditioning on stable amplification in low-rank RNNs.

@@

\begin{figure}
\vspace{-1cm}
\begin{center}
\includegraphics[scale=0.2]{figures/fig2/fig2.pdf}
\end{center}
\vspace{-1cm}
\caption{\small Emergent property inference of a stochastic stabilized supralinear network.
\textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).   
Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_b$.
Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
\textbf{B}. Responses of the deterministic smodel ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
\textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
\textbf{D}. EPI posterior of noise parameteres $\mathbf{z}$ conditioned on realistic E-population Fano factors.
The posterior predictive distribution is show on the bottom-left.
and the mode of the distribution is starred.
\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posterior.
Each gray dot is a choice of $\sigma_P$, for which a constrained mode $z^*(\sigma_P, P)$ is chosen.
The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
(Bottom) Such sensitive dimensions of the Hessian (dots) are significantly more sensitive than randomly chosen dimensions (box and whiskers).
\textbf{F}. The Fano factor of the E-population is strongly correlated with each other neuron-type population.
\textbf{G}. Mean and standard deviation (across EPI posterior) of Fano factor of each neuron-type population at each level of contrast.
}
 \label{fig:V1}
\end{figure}

@@

The slow noise of the SSSN is an an Ornstein-Uhlenbeck process
\begin{equation}
\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB,
\end{equation}
parameterized by $\sigma_\alpha$, which can be different for each neuron type,
\begin{equation}
\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top.
\end{equation} 
For this SSSN, we are interested in the parameters of slow noise that produce realistic stochastic fluctuations.
Here, we quantify this emergent property as having an excitatory population Fano factor near 1:
\begin{equation}\label{eq:EP_V1}
\begin{split}
\mathcal{X} ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 1 \\ 
 \text{Var}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  0.125^2,
\end{split}
\end{equation}
where $F_\alpha(\mathbf{x}; \mathbf{z})$ is the Fano factor of the $\alpha$-population.

@@

In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}A). 
Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$.
The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.
The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
\begin{equation}
\begin{split}
\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
\mathbf{x} &= \phi(\mathbf{u})
\end{split}
\end{equation}
with white noise of variance $0.2^2$.
The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
Here, we use EPI to determine the changes in network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ resulting in execution of rapid task switching behavior.

\begin{figure}
\begin{center}
\includegraphics[scale=0.48]{figures/fig3/fig3.pdf}
\end{center}
\caption{\small 
\textbf{A}. Rapid task switching behavioral paradigm (see text). 
\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
\textbf{C}. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice of simulations (D).
\textbf{D}. Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.  Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (optogenetic strength $\gamma=0.7$).  Shading indicates standard deviation across trials.
\textbf{E}. Difference in performance of each task during inactivation.  Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).  We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
\textbf{F}. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
}
\label{fig:SC}
\end{figure}

@@

Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
This shows that strong task representations and hemispheral dominance in the dynamics result in better execution of the Pro task.
By visualizing these four variables together by $p_A$ (Fig. \ref{fig:SC3}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
Thus, stronger hemispheral dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.

In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig. \ref{fig:SC}E) e.g. (Fig. \ref{fig:SC}D, bottom).
This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
Even though $p_P$ and $p_A$ are anticorrelated in the EPI posterior ($r=-0.15$, $p=3.68 \times 10^-12$), greater $p_P$ and $p_A$ both result in decreased resiliency to delay perturbation in the Anti task (Fig. \ref{fig:SC4}).
Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation in the Anti task (Fig. \ref{fig:SC}F)).

In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
This posterior revealed the mechanisms leading to greater accuarcy in each task as well as those increasing resiliency to perturbation in the Anti task.
Importantly, every connectivity from this inferred distribution predicts fragility and robustness of performance in the Anti and Pro tasks, respectively.
EPI allows us to conclude that since \textit{all} parameters of this model producing rapid task switching make such an experimentally verified prediction, we have a well chosen model.

@@

Here, we are interested in the scalability of EPI in number of parameters ($|\mathbf{z}|$).  We consider rank-2 RNN with N neurons of connectivity
\begin{equation}
 W = UV^\top + g\chi
 \end{equation}
 and dynamics
 \begin{equation}
 \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
 \end{equation}
where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.

@@

In neuroscience, machine learning has primarily been used to reveal structure in large-scale neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}). 
 Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
 However, these statistical models lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it. 
In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.  
In this work, we replaced this opaque procedure of parameter identification in theoretical neuroscience with emergent property inference, opening the door to careful inference in careful models of neural computation.

@@

@@

\textbf{Data availability statement}: \\
The datasets generated during and/or analysed during the current study are available from the corresponding author upon reasonable request.

@@

Consider model parameterization $\mathbf{z}$, which is a collection of scientifically interesting variables that govern the complex simulation of data $\mathbf{x}$.
For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials of the five neurons.
In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
Even so, we are not so much scientifically interested in reasoning about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
In the STG example, $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by $\mathbf{z}$.
With EPI, we learn distributions of $\mathbf{z}$ that results in an average and variance of $f(\mathbf{x}; \mathbf{z})$, denoted $\bm{\mu}$ and $\bm{\sigma}^2$.
We refer to the collection of these statistical moments as an emergent property.
Such emergent properties $\mathcal{X}$ are defined through choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), $\bm{\mu}$, and $\bm{\sigma}^2$
 \begin{equation}
\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
\end{equation}
Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and stochasticity of the data given the parameters.

In EPI, deep probability distributions are used as posterior approximations $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
 In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, .. $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
\begin{equation}
\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
\end{equation}
Such deep probability distributions embed the posterior distribution in a deep network.
Once optimized, this deep network representation has remarkably useful properties: immediate posterior sampling, and immediate probability, gradient, and Hessian evaluation at any parameter choice.

@@

 \subsubsection{Related approaches}\label{methods_related}
Inverse problems across scientific fields have been approached in a variety of ways.
The goals of such approaches are generally to a.) identify parametric sensitivities and manifolds of degeneracy with respect to some model output (structural identifiability analysis), and to reason about the parameters that are most probable to have produced a model output (statistical inference).
While much research in computational neuroscience has focused on optimizing neural architectures to process information and accomplish tasks \cite{richards2019deep}, structure in parameter space of the set of optimized solutions is rarely discussed and lacks a probabilistic treatment.

To understand sensitivity around a point, one can measure the Jacobian. One approach that scales well is EAR \cite{karlsson2012efficient}.
A popular efficient approach for systems of ODEs has been neural ODE adjoint \cite{chen2018neural} and its stochastic adaptation \cite{li2020scalable}.
Casting identifiability as a statistical estimation problem, the profile likeilhood can assess via iterated optimization while holding parameters fixed \cite{raue2009structural}.
An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point \cite{raman2017delineating}.
Global structual non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.

MCMC approaches

ABC approaches

NF LFI approaches

>>>>

EPI optimizes the weights and biases $\bm{\theta}$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. \ref{fig:STG1}).
%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
Such convergence is evaluated with a hypothesis test that the means and variances of each emergent property statistic are not different than their constrained values (see Section \ref{methods_AL_opt}). 
Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).

In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $f(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
A key advantage of EPI over established Bayesian inference is that the predictions made by the inferred distribution are constrained to produce the specified emergent property.
%Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
Equipped with this method, we may examine structure in posterior distributions or make comparisons between posteriors conditioned at different levels of the same emergent property statistic.
In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using it to investigate and produce novel insights into two prominent models in neuroscience. 
Subsequently in Section \ref{results_LRRNN}, we show EPI's superiority in parameter scalability and fidelity of the posterior predictive distribution by conditioning on stable amplification in low-rank RNNs.

@@

\begin{figure}
\vspace{-1cm}
\begin{center}
\includegraphics[scale=0.2]{figures/fig2/fig2.pdf}
\end{center}
\vspace{-1cm}
\caption{\small Emergent property inference of a stochastic stabilized supralinear network.
\textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).   
Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_b$.
Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
\textbf{B}. Responses of the deterministic model ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
\textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on realistic E-population Fano factors.
The posterior predictive distribution is show on the bottom-left.
and the mode of the distribution is starred.
\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posterior.
Each gray dot is a choice of $\sigma_P$, for which a constrained mode $z^*(\sigma_P, P)$ is chosen.
The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
(Bottom) Such sensitive dimensions of the Hessian (dots) are significantly more sensitive than randomly chosen dimensions (box and whiskers).
\textbf{F}. The Fano factor of the E-population is strongly correlated with each other neuron-type population.
\textbf{G}. Mean and standard deviation (across EPI posterior) of Fano factor of each neuron-type population at each level of contrast.
}
 \label{fig:V1}
\end{figure}

@@

The slow noise of the SSSN is an Ornstein-Uhlenbeck process
\begin{equation}
\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB,
\end{equation}
parameterized by $\sigma_\alpha$, which can be different for each neuron type,
\begin{equation}
\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top.
\end{equation} 
For this SSSN, we are interested in the parameters of slow noise that produce realistic stochastic fluctuations.
Here, we quantify this emergent property as having an excitatory population Fano factor near 1:
\begin{equation}\label{eq:EP_V1}
\begin{split}
\mathcal{X} ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 1 \\ 
 \text{Var}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  0.125^2,
\end{split}
\end{equation}
where $F_\alpha(\mathbf{x}; \mathbf{z})$ is the Fano factor of the $\alpha$-population.

@@

In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}A). 
Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$.
The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.
The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
\begin{equation}
\begin{split}
\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
\mathbf{x} &= \phi(\mathbf{u})
\end{split}
\end{equation}
with white noise of variance $0.2^2$.
The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
Here, we use EPI to determine the changes in network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ resulting in execution of rapid task switching behavior.

\begin{figure}
\begin{center}
\includegraphics[scale=0.48]{figures/fig3/fig3.pdf}
\end{center}
\caption{\small 
\textbf{A}. Rapid task switching behavioral paradigm (see text). 
\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
\textbf{C}. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice of simulations (D).
\textbf{D}. Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.  Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (optogenetic strength $\gamma=0.7$).  Shading indicates standard deviation across trials.
\textbf{E}. Difference in performance of each task during inactivation.  Inactivation level $\gamma$ scales from no inactivation (0) to full inactivation (1).  We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
\textbf{F}. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
}
\label{fig:SC}
\end{figure}

@@

Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
This shows that strong task representations and hemispheric\begin{center}
•
\end{center} dominance in the dynamics result in better execution of the Pro task.
By visualizing these four variables together by $p_A$ (Fig. \ref{fig:SC3}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
Thus, stronger hemispheric dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.

In agreement with experimental results from Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig. \ref{fig:SC}E) e.g. (Fig. \ref{fig:SC}D, bottom).
This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
Even though $p_P$ and $p_A$ are anticorrelated in the EPI posterior ($r=-0.15$, $p=3.68 \times 10^-12$), greater $p_P$ and $p_A$ both result in decreased resiliency to delay perturbation in the Anti task (Fig. \ref{fig:SC4}).
Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation in the Anti task (Fig. \ref{fig:SC}F)).

In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
This posterior revealed the mechanisms leading to greater accuracy in each task as well as those increasing resiliency to perturbation in the Anti task.
Importantly, every connectivity from this inferred distribution predicts fragility and robustness of performance in the Anti and Pro tasks, respectively.
EPI allows us to conclude that since \textit{all} parameters of this model producing rapid task switching make such an experimentally verified prediction, we have a well-chosen model.

@@

Here, we are interested in the scalability of EPI in number of parameters ($|\mathbf{z}|$).  We consider rank-2 RNN with N neurons of connectivity
\begin{equation}
 W = UV^\top + g\chi
 \end{equation}
 and dynamics
 \begin{equation}
 \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} + W\mathbf{x}
 \end{equation}
where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.

@@

In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}). 
 Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
 However, these statistical models lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it. 
In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.  
In this work, we replaced this opaque procedure of parameter identification in theoretical neuroscience with emergent property inference, opening the door to careful inference in careful models of neural computation.

@@

\textbf{TODO} *merge this point in*

While much research in computational neuroscience has focused on optimizing neural architectures to process information and accomplish tasks \cite{richards2019deep}, structure in parameter space of the set of optimized solutions is rarely discussed and lacks a probabilistic treatment.  Talk about Wyktor's work here \cite{mlynarski2020statistical}.

@@

\textbf{Data availability statement}: \\
The datasets generated during and/or analyzed during the current study are available from the corresponding author upon reasonable request.

@@

Consider model parameterization $\mathbf{z}$, which is a collection of scientifically interesting variables that govern the complex simulation of data $\mathbf{x}$.
For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials of the five neurons.
In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
Even so, we do not scientifically reason about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
In the STG example, $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by $\mathbf{z}$.
With EPI, we learn distributions of $\mathbf{z}$ that results in an average and variance of $f(\mathbf{x}; \mathbf{z})$, denoted $\bm{\mu}$ and $\bm{\sigma}^2$.
We refer to the collection of these statistical moments as an emergent property.
Such emergent properties $\mathcal{X}$ are defined through choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), $\bm{\mu}$, and $\bm{\sigma}^2$
 \begin{equation}
\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
\end{equation}
Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and stochasticity of the data given the parameters.

In EPI, deep probability distributions are used as posterior approximations $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
 In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, .. $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
\begin{equation} \label{eq:deep_transform}
\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
\end{equation}
Such deep probability distributions embed the posterior distribution in a deep network.
Once optimized, this deep network representation has remarkably useful properties: immediate posterior sampling, and immediate probability, gradient, and Hessian evaluation at any parameter choice.

@@

 \subsubsection{Related approaches}\label{methods_related}
 When Bayesian inference problems lack conjugacy, scientists use approximate inference methods like variational inference (VI) \cite{saul1998mean} and Markov chain Monte Carlo (MCMC) \cite{metropolis1953equation, hastings1970monte}. 
After optimization, variational methods return a parameterized posterior distribution, which we can analyze.
Also, the variational approximating distribution class is often chosen such that it permits fast sampling.
In contrast MCMC methods only produce samples from the approximated posterior distribution.
No parameterized distribution is estimated, and additional samples are always generated with the same sampling complexity.
Inference in models defined by systems of differential has been demonstrated with MCMC \cite{girolami2011riemann}, although this approach requires tractable likelihoods.
Advancements have leveraged structure in stochastic differential equation models to improve likelihood approximations, thus expanding the domain of applicable models \cite{golightly2011bayesian}.

Likelihood-free (or ``simulation-based") inference (LFI) \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
The most prevalent approach to LFI is approximate Bayesian computation \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
The obtained set of parameters do not have a probabilities, and further insight about the model must be gained from examination of the parameter set and their generated activity.
Methodological advances to ABC methods have come through the use of Markov chain Monte Carlo (MCMC-ABC) \cite{marjoram2003markov} and sequential Monte Carlo (SMC-ABC) \cite{sisson2007sequential} sampling techniques.
SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:LRRNN}).
Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again with a new population of initialized particles to obtain additional samples.

For scientific model analysis, we seek a posterior distribution exhibiting the properties of a well-chosen variational approximation: a parametric form conferring analytic calculations, and trivial sampling time.
For this reason, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples and have unchanging sampling rate.
EPI executes likelihood-free inference using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational posterior approximation.
The deep neural network of EPI defines the parametric form of the posterior approximation.
Furthermore, the EPI distribution is constrained to produce an emergent property.
In other words, the summary statistics of the posterior predictive distribution are fixed to have certain first and second moments.
EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
The analytic relationship between EPI and variational inference is explained in Secton \ref{methods_VI}.

 We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}).%, preprint posted simultaneously with this preprint).
We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
